---
title: "DAB-501 Group Project 01 (Manik Shakya: w0786248, Yashu Bhati: w0784198)"
output: html_document
---

Documentation of the dataset

    Author/Collaborator: David Lapp (Owner)
    Source Link : https://www.kaggle.com/datasets/johnsmith88/heart-disease-dataset
    Data (CSV): https://www.kaggle.com/datasets/johnsmith88/heart-disease-dataset/download
    

Contribution to project:
Manik and Yashu: We both contributed equally in this project. We discussed every aspect of the plot before plotting. We had a brief discussion about all the possible visualization we could make with our dataset. We each had our opinion about the plots. I helped her plotting the graphs where she needed help and she helped me to answer questions and datasets.



Loading all the needed libraries.
```{r}
library(tidyverse)
library(here)
library(ggpubr)
```

```{r}
packageVersion("tidyverse") # Version: 1.3.1
packageVersion("ggplot2") # Version: 3.3.6
packageVersion("ggpubr") # Version: 0.4.0
```


```{r}
# Read the dataset from the csv file
dataset <- read_csv("heart.csv")
```


```{r}
# Print the number of rows and columns of dataset
dim(dataset)

```


```{r}
# Get the type of dataset
typeof(dataset)
```


```{r}
# Summary of the dataset
summary(dataset)
```

```{r}
# See the structure of the data
str(dataset)
```

```{r}
# Get the first 6 rows of data
head(dataset)
```
```{r}
# Get the last 6 rows of data
tail(dataset)
```
```{r}
# Adding a label for Sex. Human readable from number to string
sex_color <- factor(dataset$sex,levels = c(0, 1), labels=c("Male","Female"))
```

```{r}
# Get the age group of people's age
age_group <- cut(dataset$age, seq(0, 100, 10))
```


Two Plots displaying the distribution of a single continuous variable.

```{r}
# Plot for continuous variable.
# Plot show the are group on x-axis and total count on y-axis.
# Distinguished by color for Male and Female
ggplot(data = dataset, mapping = aes(x = age, fill=sex_color)) + 
  geom_bar() + 
  ggtitle("Histogram plot showing age group") + 
  xlab("Age Group") + 
  ylab("Total count of age group")
```

```{r}
# Plots for continuos variable
ggplot(data=dataset, mapping = aes(chol, colour = sex_color)) + 
  geom_histogram(fill = "white",
                 bins = 15)  + 
  ggtitle("Histogram showing Cholestrol levels") + 
  xlab("Cholestrol levels") + 
  ylab("Level Count") +
  theme(legend.position = "top")
```

Two plots displaying information about a single categorical variable

```{r}
# This plot shows age group of people with ST depression induced by exercise
ggplot(data=dataset, mapping = aes(x = age, y = oldpeak, colour = sex_color)) + 
  geom_point() + 
  scale_x_discrete(labels = c("0" = "Male", "1" = "Female")) + 
  ggtitle("Plot showing ST depression level of people") + 
  xlab("Age") + 
  xlim(c(0, 100)) + 
  ylab("ST Depression Level") + 
  theme(legend.position = "top")
```

```{r}
# Plots displaying information about single categorical variable.
ggplot(data=dataset, mapping = aes(x = sex, colour = sex_color)) + 
  geom_histogram() + 
  scale_x_discrete(limits = c(0, 1), labels = c("0" = "Male", "1" = "Female")) + 
  ggtitle("Counting number of people by Gender") + 
  xlab("Sex") + 
  ylab("Count of People in Age Group")
```

One plot displaying information about both a continuous variable and a
categorical variable.

```{r}
# Plot displaying information about both a continuous and categorical variable
ggplot(data=dataset, mapping = aes(x = sex, y = age, colour = sex_color)) + 
  geom_point() + 
  scale_x_discrete(limits = c(0, 1), labels = c("0" = "Male", "1" = "Female")) + 
  xlab("Sex") + 
  ylab("Age")
```

Scattered Graph showing age group below and above their Cholestrol level.
```{r}
ggscatter(dataset, x = "age", y = "thalach", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Age", ylab = "Serum Cholesterol in mg/dl")
```


```{r}
# This plot shows age group of people with ST depression induced by exercise
ggplot(data=dataset, mapping = aes(x = age, y = oldpeak, colour = sex_color)) + 
  geom_point() + 
  scale_x_discrete(labels = c("0" = "Male", "1" = "Female")) + 
  ggtitle("Plot showing ST depression level of people") + 
  xlab("Age") + 
  xlim(c(0, 100)) + 
  ylab("ST Depression Level") + 
  theme(legend.position = "top")
```

Two plots should display information that shows a relationship between two
variables.

```{r}
# This plot shows Cholestrol and ST depression level of people distinguished by Sex
ggplot(data=dataset, mapping = aes(x = chol, y = oldpeak, colour = sex_color)) + 
  geom_point() + 
  ggtitle("Relationship between two variables") + 
  xlab("Cholestrol levels") + 
  ylab("ST depression induced by exercise") + 
  theme(legend.position = "top")
```

```{r}
# This plot shows Cholestrol and ST depression level of people distinguished by Age Group
ggplot(data=dataset, mapping = aes(x = thalach, y = oldpeak, colour = age_group)) + 
  geom_point() + 
  ggtitle("Relationship between two variables") + 
  xlab("Maximum Heart Rate Achieved") + 
  ylab("ST depression induced by exercise") + 
  theme(legend.position = "top")
```


One plot should use faceting and display information about 4 variables
```{r}
dataset$cp <- factor(dataset$cp, levels = c("0", "1", "2", "3"), 
                      labels = c("Stable angina", "Unstable angina", "Microvascular Angina", "Vasospastic or variant angina"))
dataset
ggplot(dataset) +
  geom_point(aes(x = age, y = oldpeak)) +
  facet_wrap(~ cp, nrow =2)
```


<h1>Answers to Project Questions:</h1>

<strong>Answer-1: Data visualization is important in understanding data set in the following ways:</strong>
 1) The visual context such as graphs and maps, etc gives us a clear idea of information in      a precise way. 
 2) It provides an interactive, quick and comprehend way for the human mind to understand the     data easily.
 3) The available data sets and the insights provide aid in decision analysis to decision        makers. 
 4) Moreover, Unbiased data sets without inaccuracies allows access to the right kind of information and data visualization to represent that information and keep it relevant.


<strong>Answer-2: Data visualization is important to communicating important aspects of data set in the following ways:</strong> 
 1) To communicate the information clearly and efficiently via the plots and graphs
 2) It becomes easy to identify the patterns within large data sets
 3) It helps the users to analyze the data and give proper evidences
 4) Visualization helps to understand the data sets better to measure its impact on the          business and communicates the insight visually to internal and external audiences.
 5) Moreover, this makes the data more natural for the human mind to comprehend and therefore     makes it easier to identify trends, patterns, and outliers within large data sets


<strong>Answer-3: The Role which integrity plays as an analyst when creating a data visualization for     communicating results to others are: </strong>
 1) To make even the large and/or complex data more understandable, accessible and usable
 2) To transform, improve and integrate the data depending on the business requirements          provided 
 3) To combine the dataset results across multiple sources 
 4) To deliver the data in a useful and appealing way to the audience or users
 5  To define and mention all the data variables clearly 
 6) To ensure the consistency of the data and avoid distorting the information and facts         provided in the dataset

<strong>Answer-4:</strong> More variables in data visualization increases misinterpretation of information in     datasets which result in least clarity.So we need to limit visualization to as few           variable as possible.Any Data visualization with exceeding 4 variables will be clutter       and confusion are failure of design not attributes of information result in least            effective visualization .When we exceed this number for maximum variable at the same time     in a two axis chart
    have too much statistics and intersection between correlated variable .This moderately       increases the incoherence between data.
